\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\pagecolor[rgb]{0,0,0} %black
\color[rgb]{0.75,0.75,0.75} %grey

\title{Laboration 1}
\author{Markus Svan}
\date{2021/11/22}
\begin{document}
    \maketitle
    \newpage
    \tableofcontents
    \newpage
    \section{Introduction}
    This report is about testing different sorting and search algorithms. The different sorting algorithms that are being tested in this report are the following \texttt{Bubble Sort}, \texttt{Insertion Sort}, and \texttt{QuickSort}
    The search algorithms that get tested are \texttt{Linear Search} and \texttt{Binary Search}\\
    The report will look at the complexity of each algorithm in ``big O" notation which is used to explain the expected time complexity of an algorithm based on the size of the list needing to be sorted which is noted down as \texttt{n}
    \section{Algorithm Explanation}
    \subsection{Bubble Sort}
    Bubble Sort steps through the given list of items from the lowest index to the next highest index and compares the element it is currently looking at with the element adjacent one 
    index higher, if the element at the current index is bigger than the index adjacent to it they swap places and the algorithm continues.
    After each iteration of the list it goes back to the lowest index but will then go up to the past highest index minus one.
    This continues until either there's an iteration where no swaps had to be made, meaning the list is already sorted or until the lowest index and past highest index become the same.\\
    The best case for Bubble Sort is if the list is already sorted. This results in a complexity of \texttt{O(n)} as the algorithm will step through the list once and then return since it didn't make a single swap.\\
    The worst case for Bubble Sort is if the list is already sorted in the opposite direction. This results in a complexity of \texttt{O(n²)} because the algorithm has to step through the list making a swap on every single iteration of the loop\\
    The average case for Bubble Sort is when the algorithm is given a random list. After benchmarking Bubble Sort average case through \texttt{analyze.c > bench\_bubble} the measured times somehow result in a time complexity worse than \texttt{O(n²)}.
    \subsection{Insertion Sort}
    Insertion Sort steps through the given list of items from the next lowest index to the highest index. On each step it stores the value of the current index.
    If the stored value is lower than the value one index before it'll enter a loop that goes from the current index towards the start of the list,
    constantly overwriting the new current index with the value before it until the stored value is no longer smaller than the new index and finally put the stored value on the index of the second loop.\\
    The best case for Insertion Sort is an already sorted list. This results in a complexity of \texttt{O(n)} because the algorithm will step through the list once and never have to move any values before it finishes.\\
    The worst case for Insertion Sort is an array sorted in the opposite direction. This results in a complexity of \texttt{O(n²)} due to the algorithm looping through the entire list of items and has to constantly move values during the second loop.\\
    The average case for Insertion Sort is a random list. After benchmarking Insertion Sort average case through \texttt{analyze.c > bench\_insertion} the measured times while faster than worst case still closely resemble a complexity of \texttt{O(n²)}.
    \subsection{Quick Sort}
    Quick sort selects the last element in the list as a pivot element and puts all values smaller than the pivot element to the left of it in the list and all the values bigger than the pivot element to the right of it in the list.
    The pivot element is placed correctly in the list by comparing it to every element in the list (including itself) and counting the amount of times the pivot element is bigger or equal.
    The amount of times the pivot was bigger or equal is also the pivot elements index in the list! After having placed the pivot element on the correct the algorithm then continues on the new smaller lists to the left and or right of the pivot element.
    The best case for Insertion Sort is when the pivot element is the middle element in each list section. Sadly implementing a function to make the best case list was very hard. However I can give an example list that would be best case of 9 values.
    \texttt{1 3 4 6 8 9 2 7 5} this list is a best case list as 5 is the middle element and will be picked as a pivot. And in the next recursive call of the function 2 and 7 will be the pivot element for the new respective sections.\\
    The worst case for Quick Sort is an already sorted list. The reason for this is that the pivot element will already be in the last and correct position. So there would be no splitting of the list into smaller sections. This results in a complexity of \texttt{O(N²)}.
    The average case for Quick Sort is a random list. After benchmarking quick sort average case through \texttt{analyze.c > bench\_quick} it conforms to a complexity of \texttt{O(nLog(n))}.\\
    The theoretical best case for Quick Sort is \texttt{O(nLog(n))}.
    \subsection{Linear Search}
    Linear search compares every item in the list from the lowest index to either the highest index or when the queried value is found.
    The best case for Linear Search is when the value being searched for is the first item in the list. This results in a copmplexity of \texttt{O(1)} as there is just one comparison before the value being found it is constant time in best case.\\
    The worst case for Linear Search is when the value is not in the list. This results in a complexity of \texttt{O(n)} as it steps through every single item in the list before not finding the element.\\
    The average case for Linear Search is when the value is somewhere in the list, preferably not at the start or end but closer to the middle. The complexity for Linear Search average case is less than \texttt{0(n)} but higher than \texttt{O(1)} and could be represented by roughly \texttt{O(n/2)}.
    \subsection{Binary Search}
    
    \section{Code Justification}
    \section{Experimental Setup}
    \section{Algorithm Analysis}
    \section{Problems}
    \section{Conclusion}
    
\end{document}